---
title: "Simulation exercises"
output:
  pdf_document: default
  html_notebook: default
---

# Simulation exercises

## Two girls out of six

Of all families with exactly six children, what proportion have exactly two
girls?

Assume that each child has a 50% chance of being a girl.

Have a look at the `siumulation.Rmd` notebook from the lecture for inspiration.

First simulate one family of six children:

```{r}
six_randoms <- ...
six_booleans <- six_randoms ...
number_of_trues <- ...(six_booleans)
```

Now write a `for` loop to simulate five families:

```{r}
counts <- numeric(5)
for ( ... in ...) {
  six_randoms <- ...
  six_booleans <- six_randoms ...
  number_of_trues <- ...(six_booleans)
  counts[ ... ] <- number_of_trues
}
# Show the counts
counts
```

Show how many of the values in `counts` are equal to 2:

```{r}
sum(counts ...)
```

Now you are ready for the whole thing.  Simulate 10000 families:

```{r}
n_repeats <- 10000
counts <- numeric( ...)
for ( ... in ...) {
  six_randoms <- ...
  six_booleans <- six_randoms ...
  number_of_trues <- ...(six_booleans)
  counts[ ... ] <- number_of_trues
}
# Show the first 10 counts
...(counts, ...)
```

Show the number of counts that are equal to 2:

```{r}
n_eq_2 <- sum(counts ...)
n_eq_2
```

Finally, show the proportion of counts that are equal to 2, by dividing by the number of families you simulated.

```{r}
n_eq_2 / ...
```

Now show the proportion of families with exactly six children, that have
six girls:

```{r}
n_eq_6 <- sum(counts ...)
n_eq_6 / ...
```

You will find the answer is close to `0.5 ^ 6` (0.5 to the power of six). Can you explain why?  Call your demonstrator across if you want to check your explanation.

How we get a little more difficult.

We have assumed that the probability of any one child being a girl is 50% or 0.5.  In fact, the [proportion of boys born in the UK](https://www.gov.uk/government/statistics/gender-ratios-at-birth-in-great-britain-2010-to-2014)
is 0.513.  Hence the proportion of girls is 1-0.513 = 0.487.

Use this number to do a more realistic simulation of the chances of exactly
2 girls in a family of six.

```{r}
p_girl <- 0.487
n_repeats <- 10000
counts <- numeric( ... )
for ( ... in ...) {
  six_randoms <- ...
  six_booleans <- six_randoms ...
  number_of_trues <- ...(six_booleans)
  counts[ ... ] <- number_of_trues
}
n_eq_2 <- sum(counts ...)
n_eq_2 / ...
```

Do a simulation to find what proportion of families with exactly 8 children have exactly 2 girls.

```{r}
n_repeats <- 10000
counts <- numeric( ...)
for ( ... in ...) {
    # Get inspiration from the code in the previous simulation
    ...
}
n_eq_2 <- sum(counts ...)
n_eq_2 / ...
```

## Penalty shootout technique

Jen is the penalty taker for her football team.

She's been doing this for a long time.  She thinks she normally has a 80% chance of scoring.

In the last 15 penalties she has taken, she has been trying a new technique. She scored on all 15 penalties.

How certain can she be that this would not have happened, using the old technique?

First we simulate one trial of 15 penalties:

```{r}
randoms <- runif( ... )
random_booleans <- randoms < ...
n_trues <- ...(random_booleans)
```

Run this a few times to get a feel for how likely we are to get 15 out of 15 goals (15 out of 15 TRUE values).

Now do the simulation:

```{r}
n_repeats <- 10000
counts <- numeric( ... )
for ( ... in ...) {
    randoms <- runif( ... )
    random_booleans <- randoms < ...
    n_trues <- ...(random_booleans)
    counts[ ... ] <- ...
}
n_eq_15 <- sum(counts ...)
n_eq_15 / ...
```

What do you think?  Is 15 / 15 a crazy outcome assuming Jen's chances are exactly the same as before, or should we conclude that something has changed?

# Burkitt's lymphoma drug trial

Burkitt's lymphoma is a cancer of the human lymph system.

At the time I wrote this, you can find the following on the [Wikipedia page for Burkitt's Lymphoma](https://en.wikipedia.org/wiki/Burkitt%27s_lymphoma#Prognosis).

> The overall cure rate for Burkitt's lymphoma in developed countries is
> about 90%, but worse in low-income countries. Burkitt's lymphoma is
> uncommon in adults, where it has a worse prognosis {%cite
> molyneux2012burkitt %}
>
> In 2006, treatment with dose-adjusted EPOCH with Rituximab showed
> promising initial results in a small series of patients (n=17), with
> a 100% response rate, and 100% overall survival and progression-free
> survival at 28 months (median follow-up) {%cite dunleavy2006novel %}

How likely is it that these study results, or better, could have come
about by chance?

* Your model is that the EPOCH study was, in fact, no more effective than any other standard therapy, and that they were just lucky, in finding that 100% of their patients were cured.
* You are going to do a simulation, using this model.  In this simulation, you will make 17 simulated patients, each with a 90% chance of being cured.  Then count how many of the 17 patients were cured.
* Do this many times, to see how likely you are to get the value 17 / 17 (100%).

```{r}
n_repeats <- 1000
counts <- ...
for ( ... in ... ) {
    # You will need several lines here
    ...
    counts[ ... ] <- ...
}
n_eq_17 <- ...
# Show the proportion
...
```
